Multi-target detection with application to cryo-electron microscopy

### 2019-03-05 inverse_problems/inverse_problems.tex
section 5.2

For the experiment depicted in Figure~\ref{fig:1Dheterosignals}, we fix $K = 3$ signals of length $L = 21$: see the three red signals in the first column. Following~\eqref{eq:model}, we generate an observation $y$ of length $12.3 \cdot 10^9$. Each of the three signals appears, respectively (and approximately), $30.0 \cdot 10^6$, $20.0 \cdot 10^6$ and $10.0 \cdot 10^6$ times in $y$ for a total of exactly $60 \cdot 10^6$ occurrences, such that at least $L-1$ zeros separate any two occurrences of any signals according to~\eqref{eq:spacing}. This is done by randomly selecting $60 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint and locations picked so far. For each placement, one of the three signals is picked at random according to the proportions $\pi = (1/2, 1/3, 1/6)$. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. The resulting signal-to-noise ratio of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
is about 1/9.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.7\linewidth]{KLXP/XP1}
	\caption{In the $N \to \infty$ regime (access to exact autocorrelations, excluding biased entries) and with known uniform densities, it seems $K$ up to $\sqrt{L}$ (red curve) i.i.d.\ Gaussian signals of length $L$ can be recovered from the known moments. CPU time in seconds. Strictly above red dots, recovery is impossible because the number of unknowns exceeds the number of computed moments. Similar to~\cite[Fig.~4.1]{boumal2017heterogeneous}, this experiment suggests a possible statistical-computational gap.}
	\label{fig:KLXP}
\end{figure}

In this second experiment, presented in Figure~\ref{fig:KLXP}, we investigate how many distinct signals ${x_1, \ldots, x_K}$ can be estimated from mixed autocorrelations~\eqref{eq:mixedautocorr}. In order to do so, we consider a setup where the mixed autocorrelations are known perfectly. This corresponds to the limit of an infinitely long observation $y$ with fixed density $\gamma$ and fixed noise level (that may be arbitrarily high). The specific value of $\sigma$ is immaterial since we only consider autocorrelations that are unaffected by noise bias. Furthermore, we assume uniform occurrence distribution $\pi$ (known to the algorithm), and known density $\gamma$ (which is then irrelevant as it only induces a global scaling of the autocorrelations of $y$).

@@

Following~\cite{boumal2017heterogeneous}, overlaid on the panels we trace the red curve $K = \sqrt{L}$ as well as red dots which are computed from the considerations in Section~\ref{sec:heterogeneity} (adapted to the fact that $\pi$ and $\gamma$ are known). We observe that strictly above the red dots the optimization problem appears to be easy to solve (despite non-convexity), yet, as predicted, the corresponding estimators are not informative since there is not enough information in the computed quantities compared to the number of parameters. On the other hand, below the (empirical) red curve, the optimization problem is sometimes solved to optimality (although it may take more than one random initialization to get one successful run), and the corresponding estimators are accurate. In between the red curve and the red dots, the optimization problem appears to be particularly challenging: we essentially never produce a global optimum, hence we also do not have an estimator. This experiment suggests a possible computational-statistical gap in the area between the red curve and the red dots, where it is possible that the signals could be estimated, but perhaps not with a computationally efficient procedure.
Similar results were observed for the multi-reference problem~\cite{boumal2017heterogeneous,weinthesis,bandeira2017estimation}.

>>>>

For the experiment depicted in Figure~\ref{fig:1Dheterosignals}, we fix $K = 3$ signals of length $L = 21$: see the three red signals in the first column. Following~\eqref{eq:model}, we generate an observation $y$ of length $12.3 \cdot 10^9$. Each of the three signals appears, respectively (and approximately), $30.0 \cdot 10^6$, $20.0 \cdot 10^6$ and $10.0 \cdot 10^6$ times in $y$ for a total of exactly $60 \cdot 10^6$ occurrences, such that at least $L-1$ zeros separate any two occurrences of any signals according to~\eqref{eq:spacing}. This is done by randomly selecting $60 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint~\eqref{eq:spacing} and locations picked so far. For each placement, one of the three signals is picked at random according to the proportions $\pi = (1/2, 1/3, 1/6)$. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. The resulting signal-to-noise ratio of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
is about 1/9.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.7\linewidth]{KLXP/XP1}
	\caption{In the $N \to \infty$ regime (access to exact autocorrelations, excluding biased entries) and with known uniform densities, it seems that $K$ up to $\sqrt{L}$ (red curve) i.i.d.\ Gaussian signals of length $L$ can be recovered from the known moments. CPU time is in seconds. Strictly above red dots, recovery is impossible because the number of unknowns exceeds the number of computed autocorrelation; see Section~\ref{sec:heterogeneity}. Similarly to~\cite[Fig.~4.1]{boumal2017heterogeneous}, this experiment suggests a possible statistical-computational gap.}
	\label{fig:KLXP}
\end{figure}

In this second experiment, presented in Figure~\ref{fig:KLXP}, we investigate how many distinct signals ${x_1, \ldots, x_K}$ can be estimated from mixed autocorrelations~\eqref{eq:mixedautocorr}. In order to do so, we consider a setup where the mixed autocorrelations are known perfectly. This corresponds to the limit of an infinitely long observation $y$ with fixed density $\gamma$ and fixed noise level (that may be arbitrarily high). The specific value of $\sigma$ is immaterial since we only consider autocorrelations that are unaffected by noise bias. Furthermore, we assume uniform occurrence distribution $\gamma_k$ (known to the algorithm), and known density $\gamma$ (which is then irrelevant as it only induces a global scaling of the autocorrelations of $y$).

@@

Following~\cite{boumal2017heterogeneous}, overlaid on the panels we trace the red curve $K = \sqrt{L}$ as well as red dots which are computed from the considerations in Section~\ref{sec:heterogeneity} (adapted to the fact that the $\gamma_k$'s are known). We observe that strictly above the red dots the optimization problem appears to be easy to solve (despite non-convexity), yet, as predicted, the corresponding estimators are not informative since there is not enough information in the computed quantities compared to the number of parameters. On the other hand, below the (empirical) red curve, the optimization problem is sometimes solved to optimality (although it may take more than one random initialization to get one successful run), and the corresponding estimators are accurate. In between the red curve and the red dots, the optimization problem appears to be particularly challenging: we essentially never produce a global optimum, hence we also do not have an estimator. This experiment suggests a possible computational-statistical gap in the area between the red curve and the red dots, where it is possible that the signals could be estimated, but perhaps not with a computationally efficient procedure.
Similar results were observed for the multi-reference problem~\cite{boumal2017heterogeneous,weinthesis,bandeira2017estimation}.
